Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2017-05-01 apj.tex
complete new version that sets length of R vector

\begin{abstract}
\color{orange}
Through empirical modeling, the peak absolute magnitude of a Type~Ia supernova (SN~Ia) can be accurately determined 
based on its observed signal, making SNe~Ia  excellent distance indicators.  Improved modeling of SN~Ia absolute magnitudes can account for
further physical
diversity that is expected but not included
in current models, and so consequently can lead to more precise per-object distances with tighter systematic uncertainties.  In this article, we present the
results of    
\color{black}
an empirical model for SN~Ia peak magnitudes with two color parameters and dependence on the equivalent widths of CaII, SiII, and SiII velocity
as applied to the supernova sample of the Nearby Supernova Factory.  The peak magnitudes in synthetic
broadband photometry and their colors are found to be 
dependent on the spectral equivalent widths and the color parameters, to better than 
%-----
$(1-5\times10^{-5})$
%------
confidence
\color{orange}
based on the posterior.
\color{black}
The color changes allowed by the model are almost parallel to the color changes allowed by the extinction model of  \citet{1999PASP..111...63F} with floating
total-to-selective extinction parameter $R^F$.  The deduced color excess
is consistent with an effective $R^F = 2.44$.
The derived dust curve is able to reproduce the optical colors of the highly reddened SN~2014J better than the \citet{1999PASP..111...63F} dust model.
Inspired that our  two-parameter model, with no conditioning that would steer it toward a dust model, produces behavior similar to that expected from Galactic dust, we extend  the model to three color parameters.
In this new case we obtain consistent results for
all the parameters in common,
plus a detection of a third parameter at
%--- table.py
$(1-5\times10^{-4})$
%-----
confidence;
this new parameter represents supernova color diversity
beyond that tracked by the spectral features and two-parameter dust models.
%A direct incorporation of dust models into our formalism is impossible. An indirect incorporation of dust
%yields in constant-$R^F$ model $R^F \sim 3$, but the implementation is sensitive to systematic uncertainties that arise
%from the need to specify the colors of an unextincted supernova and its use of absolute extinction, neither of which is used in our original model.
\end{abstract}

@@

\color{orange}
We impose
\begin{equation}
\langle \Delta \rangle=0, \langle k_0 \rangle=0, \langle k_1 \rangle=0.
\end{equation}
These conditions specify the definition of absolute magnitude and zero color relative to which the color excess is measured:
in our linear model these values are unconstrained.
\color{orange}
We impose 
\begin{equation}
\gamma^0_U > 0, \gamma^1_U < 0.
\end{equation}
to exclude degenerate posterior space
associated
with the $\pm$ sign, i.e.\ simultaneous sign flips of
$\gamma^0$--$k_0$ and $\gamma^1$--$k_1$
(when running our fits, $k_0$ and $k_1$ are set to
identical initial conditions, so the difference in the signs of $\gamma^0_U$ and $\gamma^1_U$ distinguishes the two.)

@@

\color{orange}
For four selected colors, a plot of the
observed versus deduced colors and the histogram of their differences
\color{black}
are shown in Figure~\ref{residual:fig}. 
There is no evidence of a catastrophic fit nor of extreme
outlying data.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/residual.pdf} 
            \caption{
            \color{orange}
            Observed versus deduced colors for four selected colors.  Plotted on the right for each color
is the histogram of the difference between the observed and deduced colors.
\color{black}
            \label{residual:fig}}
\end{figure}

@@

The histogram of the medians of the grey offsets, $\Delta$, for all supernovae is shown in Figure~\ref{hist:fig}.  The distribution is non-Gaussian, 
has a standard deviation of
%-----
$0.10$
%-----
mag, and a broad tail in the positive (fainter) direction. 
\color{orange}
Of the three extreme outliers, the supernovae with largest and third largest $\Delta$'s have  heliocentric redshifts of 
 $0.0015$ and $0.0085$ respectively, and so have peculiar velocities that can cause a significant magnitude offset.
\color{black}
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/Delta_hist.pdf} 
   \caption{Normalized ideogram and histogram of the medians of the grey offset $\Delta$. 
   \label{hist:fig}}
\end{figure}

Non-trivial residual magnitude dispersions are captured in $C_c$.  The diagonal elements are captured by the $\sigma$ parameters;
the residual intrinsic dispersion ranges from
$\sim 0.01$ to 0.06 mag, significantly smaller
than the dispersion in $\Delta$.
\color{orange}
Of particular interest within $C_c$ is intrinsic color dispersion.  The intrinsic dispersions in $V$ and colors with respect to $V$ are
\begin{equation}
\sigma(V,U-V,B-V,V-R,V-I)=
%\begin{pmatrix}
%\begin{array}{rrrrr}
\left(
0.021^{+0.004}_{-0.005} ,
0.067^{+0.010}_{-0.010},
0.034^{+0.005}_{-0.005},
0.023^{+0.002}_{-0.002},
0.053^{+0.004}_{-0.004}
\right)
%\end{array}
% \end{pmatrix} 
 \text{mag}.
 \label{sig_intrinsic:eqn}
 \end{equation}
The correlations between these parameters are
\begin{equation}
Cor(V,U-V,B-V,V-R,V-I)=
\begin{pmatrix}
\begin{array}{rrrrr}
1 & -0.48^{+0.20}_{-0.18} & -0.34^{+0.21}_{-0.22} & 0.88^{+0.10}_{-0.25} & 0.60^{+0.11}_{-0.19} \\
-0.48^{+0.20}_{-0.18} & 1 & 0.60^{+0.10}_{-0.13} & -0.38^{+0.18}_{-0.17} & -0.33^{+0.15}_{-0.14} \\
-0.34^{+0.21}_{-0.22} & 0.60^{+0.10}_{-0.13} & 1 & -0.21^{+0.17}_{-0.17} & 0.08^{+0.14}_{-0.15} \\
0.88^{+0.10}_{-0.25} & -0.38^{+0.18}_{-0.17} & -0.21^{+0.17}_{-0.17} & 1 & 0.65^{+0.06}_{-0.07} \\
0.60^{+0.11}_{-0.19} & -0.33^{+0.15}_{-0.14} & 0.08^{+0.14}_{-0.15} & 0.65^{+0.06}_{-0.07} & 1 \\
\end{array}
 \end{pmatrix}.
 \label{cor_intrinsic:eqn}
 \end{equation}
\color{black}
%
%Given that the 
%off-diagonal elements of $C_c$ are parameterized by the Cholesky factors of a correlation matrix rather than the matrix elements themselves,
%it is not straightforward to present correlation matrix confidences:
%the average over Monte Carlo links of the
%Cholesky factors will not  generally yield a correlation matrix.  
%To characterize a typical posterior draw of $C_c$ we use the matrix that is the mean of all covariance realizations in the
%chain, element by element.
%For $UBVRI$ the matrix is
%\begin{equation}
%C_c(U,B,V,R,I)=
%\begin{pmatrix}
%\begin{array}{rrrrr}
%0.0038 & 0.0010 & -0.0002 & -0.0000 & 0.0003 \\
%0.0010 & 0.0011 & 0.0002 & -0.0000 & -0.0006 \\
%-0.0002 & 0.0002 & 0.0004 & 0.0000 & -0.0002 \\
%-0.0000 & -0.0000 & 0.0000 & 0.0002 & 0.0002 \\
%0.0003 & -0.0006 & -0.0002 & 0.0002 & 0.0020
%\end{array}
% \end{pmatrix} \text{mag}^2.
% \label{mag_cov:eqn}
% \end{equation}
%The average correlation matrix is
%\begin{equation}
%Cor(U,B,V,R,I)=
%\begin{pmatrix}
%\begin{array}{rrrrr}
%1.00 & 0.42 & -0.18 & -0.05 & 0.12 \\
%0.42 & 1.00 & 0.26 & -0.09 & -0.39 \\
%-0.18 & 0.26 & 1.00 & 0.09 & -0.22 \\
%-0.05 & -0.09 & 0.09 & 1.00 & 0.27 \\
%0.12 & -0.39 & -0.22 & 0.27 & 1.00
%\end{array}
% \end{pmatrix}.
% \end{equation}
%The  covariance of the colors $U-V$, $B-V$, $V-R$, and $V-I$ is
%expressed as the standard deviations and
%correlation matrix
% \begin{equation}
% \sigma(U-V, B-V, V-R, V-I)=
% \begin{pmatrix}
%0.068 & 0.034 & 0.023 & 0.053
%  \end{pmatrix} \text{mag},
% \label{color_sd:eqn}
% \end{equation}
% \begin{equation}
% Cor(U-V, B-V, V-R, V-I)=
%\begin{pmatrix}
%\begin{array}{rrrr}
%1.000 & 0.602 & -0.353 & -0.317 \\
%0.602 & 1.000 & -0.197 & 0.082 \\
%-0.353 & -0.197 & 1.000 & 0.647 \\
%-0.317 & 0.082 & 0.647 & 1.000
%\end{array}
%\end{pmatrix}.
%  \label{color_cor:eqn}
% \end{equation}
These color variances are smaller than those derived by \citet{2003A&A...404..901N} or \citet{2007ApJ...659..122J}, for cases where direct comparison can be made.
\color{orange}
%model11_rubin.py
Having no intrinsic color dispersion is disfavored, and not allowing for them can have important consequences.  Running an analysis assuming no intrinsic dispersion (replacing the probabilistic intrinsic magnitudes in Eqn.~\ref{ewsiv:eqn}
with a strict equality), we find significant shifts in the inferred slopes and smaller statistical uncertainties.
\color{black}

@@

The Pearson correlation coefficients for $\Delta$, $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $E_{\gamma^0}(B-V)$, and $E_{\gamma^0}(B-V)$ are given in the matrix
\color{orange}
\begin{multline}
Cor(\Delta, EW_{Ca}, EW_{Si}, \lambda_{Si}, E_{\gamma^0}(B-V), E_{\gamma^1}(B-V)) =\\
\begin{pmatrix}
\begin{array}{rrrrrr}
1 & 0.00^{+0.06}_{-0.06} & -0.07^{+0.07}_{-0.06} & -0.04^{+0.06}_{-0.06} & 0.10^{+0.07}_{-0.07} & -0.00^{+0.08}_{-0.07} \\
0.00^{+0.06}_{-0.06} & 1 & 0.10^{+0.03}_{-0.03} & -0.26^{+0.03}_{-0.03} & -0.07^{+0.06}_{-0.06} & -0.01^{+0.08}_{-0.08} \\
-0.07^{+0.07}_{-0.06} & 0.10^{+0.03}_{-0.03} & 1 & -0.14^{+0.04}_{-0.03} & -0.17^{+0.05}_{-0.05} & -0.04^{+0.08}_{-0.07} \\
-0.04^{+0.06}_{-0.06} & -0.26^{+0.03}_{-0.03} & -0.14^{+0.04}_{-0.03} & 1 & 0.03^{+0.05}_{-0.06} & 0.03^{+0.08}_{-0.08} \\
0.10^{+0.07}_{-0.07} & -0.07^{+0.06}_{-0.06} & -0.17^{+0.05}_{-0.05} & 0.03^{+0.05}_{-0.06} & 1 & 0.06^{+0.08}_{-0.08} \\
-0.00^{+0.08}_{-0.07} & -0.01^{+0.08}_{-0.08} & -0.04^{+0.08}_{-0.07} & 0.03^{+0.08}_{-0.08} & 0.06^{+0.08}_{-0.08} & 1 \\
\end{array}
\end{pmatrix}.
\end{multline}
\color{black}

We find weak correlations between our color-excess parameters and the input features; 
\color{orange}
only the $ -0.17 \pm 0.05$ between $E_{\gamma^0}(B-V)$ and $EW_{Si}$ has significance much larger than $1 \sigma$.
\color{black}
Recall that $EW_{Si}$ is correlated with light-curve shape, which is correlated with host-galaxy (including dust) properties 
\citep{2000AJ....120.1479H, 2003MNRAS.340.1057S}.
An extensive discussion of the correlations between the spectral features of the SNfactory data set can be found in \citet{chotard:thesis}
and \citet{leget:thesis}.

@@

\color{orange}
For brevity, we do not present here the analysis of simulated data used to validate and quantify bias in our results.
The simulation tests in \S\ref{model2:sec} are representative for both Model~1 of this Section and Model~2 of the subsequent section.
\color{black}

@@

%\color{orange}

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.95in]{output11/plane0.pdf}
   \includegraphics[width=2.95in]{output11/plane1.pdf}
%   \includegraphics[width=2.95in]{output11/plane0BVR.pdf}
%   \includegraphics[width=2.95in]{output11/plane1BVR.pdf}
   \caption{
   \color{orange}
   Two perspectives (left and right columns) of the $UVI$-space
of the vectors corresponding to $\gamma^i_X/(\gamma^i_B-\gamma^i_V)$ of our model (solid lines),
   and $a(X)$, $b(X)$ of the \citetalias{1999PASP..111...63F} model (dashed lines).  The combination $a(X)+b(X)/2.44$ is shown in the dotted red
   line: it is almost perfectly superimposed on $\gamma^0_X/(\gamma^0_B-\gamma^0_V)$.   The combination $-7.31a(X)+b(X)$ is shown in the dotted blue
   line: it is
   almost  perfectly superimposed on $\gamma^1_X/(\gamma^1_B-\gamma^1_V)$ in $UVI$.  
   \label{plane:fig}}
\end{figure}

@@

\color{orange}
The preferred values for $\kappa$ can be determined from the matrix elements of Eqn.~\ref{model11trans:eqn}: $\kappa^{-1}_1=2.82/1.15 =2.44$, $\kappa_2=-5.27/0.72=-7.31$.
The directions of the $\gamma$ vectors, and  the expectations from  \citetalias{1999PASP..111...63F} with $R=\kappa^{-1}$, are shown in 
Figure~\ref{plane:fig}.  There is almost perfect alignment between $\gamma^0$ and $R=2.44$ dust.
 From the Monte Carlo chains, we find a 68\% credible interval for $\kappa^{-1}_1$ of 2.26--2.63.

@@

\color{orange}
As a check of our analysis implementation, we generate ten simulated datasets with the same numbers and uncertainties as our data, based on a
SALT2 $x_1=0$, $c=0$ template and  \citetalias{1999PASP..111...63F} dust.  The supernovae
are assigned $E^F(B-V)$ parameters equal to the $B-V$ color excesses determined in our sample
(making an approximative correspondence between our observed top-hat filter colors and the 
model parameter $E^F(B-V)$) , and $R^F$ values drawn from a log-normal
distribution with a mean of $\ln{2.5}$ and 0.05 standard deviation.  The $\alpha$, $\beta$, and $\eta$ parameters are set to zero, so that dust and
measurement uncertainty are the only sources of variation in magnitudes.  We stack the results of the 10 simulations: the input $R^F=2.5$ is centered
within the 68\% credible interval for $\kappa^{-1}_1$ of 2.38--2.69.

@@

Given the association of  our parameter $\gamma^0_V k_0$ with changes in $A^F_V$ of an $R^F =\kappa^{-1}_1$ dust model up to an additive offset,
the former's distribution can be compared to the expectation for the latter.
Figure~\ref{k0_med:fig} shows the histogram of
median values of $\gamma^0_V k_0$.
The distribution is non-Gaussian, having a sharp rise in the blue and an extended tail in the red.  This is consistent
with the expectation of dust considering the spatial distribution of supernovae within galaxies and the distribution of galaxy orientations with respect to the observer \citep{1998ApJ...502..177H}.  This distribution is given by the data, in that it is independent of any knowledge or prior
for the distribution of dust extinction the SN population is expected to suffer.

Figure~\ref{k0_med:fig} also shows the distribution of $\gamma^1_V k_1$, which can be attributed to 
extinction corrections due to deviations away from the canonical $\kappa_1$ value.   The distribution has a long
negative tail and is peaked at its positive edge, and by construction has a mean of zero.
The implication is that while the bulk of supernovae have positive values associated with $R>1/\kappa_1$, there is a tail
of negative values associated with $R<1/\kappa_1$.  The plot of   $\gamma^0_V k_0$ versus $\gamma^1_V k_1$  shown in Figure~\ref{kk:fig} shows that
the positive edge in $\gamma^1_V k_1$ persists over the full range of  $\gamma^0_V k_0$
%\color{orange}
%While at low  $\gamma^0_V k_0$
%\color{black}
%On the other hand, lower values of $\gamma^1_V k_1$
%occur at larger values of  $\gamma^0_V k_0$.
These findings are consistent with previous results:
\citet{2014ApJ...789...32B, 2015MNRAS.453.3300A} deduce a wide range of dust behavior $1.5<R^F<3$ encountered by the SN~Ia population.
%\citet{2011ApJ...729...55F} and the above authors
%find that supernovae with large extinction 
%preferentially exhibit low $R^F$.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{output11/gamma0_med.pdf}
   \includegraphics[width=2.8in]{output11/gamma1_med.pdf}
      \caption{
      \color{orange} Normalized ideogram and histogram 
      \color{black}
      of
median values of Left: $\gamma^0_V k_0$, which is associated with the $A^F_V$ of an $R^F \sim 1/\kappa_1$ dust model; Right: $\gamma^1_V k_1$,
which  is associated with extinction corrections due to deviations away from the canonical $\kappa_1$ value.
   \label{k0_med:fig}}
\end{figure}

>>>>

\begin{abstract}
\color{purple}
Through empirical modeling, the peak absolute magnitude of a Type~Ia supernova (SN~Ia) can be accurately determined 
based on its observed signal, making SNe~Ia  excellent distance indicators.  Improved modeling of SN~Ia absolute magnitudes can account for
further physical
diversity that is expected but not included
in current models, and so consequently can lead to more precise per-object distances with tighter systematic uncertainties.  In this article, we present the
results of    
\color{black}
an empirical model for SN~Ia peak magnitudes with two color parameters and dependence on the equivalent widths of CaII, SiII, and SiII velocity
as applied to the supernova sample of the Nearby Supernova Factory.  The peak magnitudes in synthetic
broadband photometry and their colors are found to be 
dependent on the spectral equivalent widths and the color parameters, to better than 
%-----
$(1-5\times10^{-5})$
%------
confidence.
The color changes allowed by the model are almost parallel to the color changes allowed by the extinction model of  \citet{1999PASP..111...63F} with floating
total-to-selective extinction parameter $R^F$.  The deduced color excess
is consistent with an effective $R^F = 2.44$.
The derived dust curve is able to reproduce the optical colors of the highly reddened SN~2014J better than the \citet{1999PASP..111...63F} dust model.
Inspired that our  two-parameter model, with no conditioning that would steer it toward a dust model, produces behavior similar to that expected from Galactic dust, we extend  the model to three color parameters.
In this new case we obtain consistent results for
all the parameters in common,
plus a detection of a third parameter at
%--- table.py
$(1-5\times10^{-4})$
%-----
confidence;
this new parameter represents supernova color diversity
beyond that tracked by the spectral features and two-parameter dust models.
%A direct incorporation of dust models into our formalism is impossible. An indirect incorporation of dust
%yields in constant-$R^F$ model $R^F \sim 3$, but the implementation is sensitive to systematic uncertainties that arise
%from the need to specify the colors of an unextincted supernova and its use of absolute extinction, neither of which is used in our original model.
\end{abstract}

@@

\color{purple}
We impose
\begin{equation}
\langle \Delta \rangle=0, \langle k_0 \rangle=0, \langle k_1 \rangle=0.
\end{equation}
These conditions specify the definition of absolute magnitude and zero color relative to which the color excess is measured:
in our linear model these values are unconstrained.
\color{purple}
We impose 
\begin{equation}
\gamma^0_U > 0, \gamma^1_U < 0.
\end{equation}
to exclude degenerate posterior space
associated
with the $\pm$ sign, i.e.\ simultaneous sign flips of
$\gamma^0$--$k_0$ and $\gamma^1$--$k_1$
(when running our fits, $k_0$ and $k_1$ are set to
identical initial conditions, so the difference in the signs of $\gamma^0_U$ and $\gamma^1_U$ distinguishes the two.)

@@

\color{purple}
For four selected colors, a plot of the
observed versus deduced intrinsic colors and the histogram of their differences
\color{black}
are shown in Figure~\ref{residual:fig}. 
There is no evidence of a catastrophic fit nor of extreme
outlying data.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/residual.pdf} 
            \caption{
            \color{purple}
            Observed versus deduced intrinsic colors for four selected colors.  Plotted on the right for each color
is the histogram of the difference between the observed and intrinsic colors.
\color{black}
            \label{residual:fig}}
\end{figure}

@@

\color{purple}
The fit $\Delta$'s are highly correlated between supernovae.  To aid in visualization, we remove much of this
correlation by subtracting out the $\Delta|_0$ of  an arbitrary supernova.
The histogram of the medians of these relative  grey offsets $\Delta-\Delta|_0$,
is shown in Figure~\ref{hist:fig}.  The distribution
has a standard deviation of
%-----
$0.08$
%-----
mag, and a tail in the positive (fainter) direction. 
Of the three extreme outliers, the supernovae with largest and third largest $\Delta$'s have  heliocentric redshifts of 
 $0.0015$ and $0.0085$ respectively, and so have peculiar velocities that can cause a significant magnitude offset.
\color{black}
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/deltaDelta_hist.pdf} 
   \caption{
   \color{purple}
   Normalized ideogram and histogram of the medians of the grey offset relative to one arbitrary supernova, $\Delta-\Delta|_0$.  
   \color{black}
   \label{hist:fig}}
\end{figure}

Non-trivial residual magnitude dispersions are captured in $C_c$.  The diagonal elements are captured by the $\sigma$ parameters;
the residual intrinsic dispersion ranges from
$\sim 0.01$ to 0.06 mag, significantly smaller
than the dispersion in $\Delta$.
\color{purple}
Of particular interest within $C_c$ is intrinsic color dispersion.  The intrinsic dispersions in $V$ and colors with respect to $V$ are
\begin{equation}
\sigma(V,U-V,B-V,V-R,V-I)=
%\begin{pmatrix}
%\begin{array}{rrrrr}
\left(
0.021^{+0.004}_{-0.005} ,
0.067^{+0.010}_{-0.010},
0.034^{+0.005}_{-0.005},
0.023^{+0.002}_{-0.002},
0.053^{+0.004}_{-0.004}
\right)
%\end{array}
% \end{pmatrix} 
 \text{mag}.
 \label{sig_intrinsic:eqn}
 \end{equation}
The correlations between these parameters are
\begin{equation}
Cor(V,U-V,B-V,V-R,V-I)=
\begin{pmatrix}
\begin{array}{rrrrr}
1 & -0.48^{+0.20}_{-0.18} & -0.34^{+0.21}_{-0.22} & 0.88^{+0.10}_{-0.25} & 0.60^{+0.11}_{-0.19} \\
-0.48^{+0.20}_{-0.18} & 1 & 0.60^{+0.10}_{-0.13} & -0.38^{+0.18}_{-0.17} & -0.33^{+0.15}_{-0.14} \\
-0.34^{+0.21}_{-0.22} & 0.60^{+0.10}_{-0.13} & 1 & -0.21^{+0.17}_{-0.17} & 0.08^{+0.14}_{-0.15} \\
0.88^{+0.10}_{-0.25} & -0.38^{+0.18}_{-0.17} & -0.21^{+0.17}_{-0.17} & 1 & 0.65^{+0.06}_{-0.07} \\
0.60^{+0.11}_{-0.19} & -0.33^{+0.15}_{-0.14} & 0.08^{+0.14}_{-0.15} & 0.65^{+0.06}_{-0.07} & 1 \\
\end{array}
 \end{pmatrix}.
 \label{cor_intrinsic:eqn}
 \end{equation}
\color{black}
%
%Given that the 
%off-diagonal elements of $C_c$ are parameterized by the Cholesky factors of a correlation matrix rather than the matrix elements themselves,
%it is not straightforward to present correlation matrix confidences:
%the average over Monte Carlo links of the
%Cholesky factors will not  generally yield a correlation matrix.  
%To characterize a typical posterior draw of $C_c$ we use the matrix that is the mean of all covariance realizations in the
%chain, element by element.
%For $UBVRI$ the matrix is
%\begin{equation}
%C_c(U,B,V,R,I)=
%\begin{pmatrix}
%\begin{array}{rrrrr}
%0.0038 & 0.0010 & -0.0002 & -0.0000 & 0.0003 \\
%0.0010 & 0.0011 & 0.0002 & -0.0000 & -0.0006 \\
%-0.0002 & 0.0002 & 0.0004 & 0.0000 & -0.0002 \\
%-0.0000 & -0.0000 & 0.0000 & 0.0002 & 0.0002 \\
%0.0003 & -0.0006 & -0.0002 & 0.0002 & 0.0020
%\end{array}
% \end{pmatrix} \text{mag}^2.
% \label{mag_cov:eqn}
% \end{equation}
%The average correlation matrix is
%\begin{equation}
%Cor(U,B,V,R,I)=
%\begin{pmatrix}
%\begin{array}{rrrrr}
%1.00 & 0.42 & -0.18 & -0.05 & 0.12 \\
%0.42 & 1.00 & 0.26 & -0.09 & -0.39 \\
%-0.18 & 0.26 & 1.00 & 0.09 & -0.22 \\
%-0.05 & -0.09 & 0.09 & 1.00 & 0.27 \\
%0.12 & -0.39 & -0.22 & 0.27 & 1.00
%\end{array}
% \end{pmatrix}.
% \end{equation}
%The  covariance of the colors $U-V$, $B-V$, $V-R$, and $V-I$ is
%expressed as the standard deviations and
%correlation matrix
% \begin{equation}
% \sigma(U-V, B-V, V-R, V-I)=
% \begin{pmatrix}
%0.068 & 0.034 & 0.023 & 0.053
%  \end{pmatrix} \text{mag},
% \label{color_sd:eqn}
% \end{equation}
% \begin{equation}
% Cor(U-V, B-V, V-R, V-I)=
%\begin{pmatrix}
%\begin{array}{rrrr}
%1.000 & 0.602 & -0.353 & -0.317 \\
%0.602 & 1.000 & -0.197 & 0.082 \\
%-0.353 & -0.197 & 1.000 & 0.647 \\
%-0.317 & 0.082 & 0.647 & 1.000
%\end{array}
%\end{pmatrix}.
%  \label{color_cor:eqn}
% \end{equation}
These color variances are smaller than those derived by \citet{2003A&A...404..901N} or \citet{2007ApJ...659..122J}, for cases where direct comparison can be made.
\color{purple}
%model11_rubin.py
Having no intrinsic color dispersion is disfavored, and not allowing for them can have important consequences.  Running an analysis assuming no intrinsic dispersion (replacing the probabilistic intrinsic magnitudes in Eqn.~\ref{ewsiv:eqn}
with a strict equality), we find significant shifts in the inferred slopes and smaller statistical uncertainties.
\color{black}

@@

The Pearson correlation coefficients for $\Delta$, $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $E_{\gamma^0}(B-V)$, and $E_{\gamma^0}(B-V)$ are given in the matrix
\color{purple}
\begin{multline}
Cor(\Delta, EW_{Ca}, EW_{Si}, \lambda_{Si}, E_{\gamma^0}(B-V), E_{\gamma^1}(B-V)) =\\
\begin{pmatrix}
\begin{array}{rrrrrr}
1 & 0.00^{+0.06}_{-0.06} & -0.07^{+0.07}_{-0.06} & -0.04^{+0.06}_{-0.06} & 0.10^{+0.07}_{-0.07} & -0.00^{+0.08}_{-0.07} \\
0.00^{+0.06}_{-0.06} & 1 & 0.10^{+0.03}_{-0.03} & -0.26^{+0.03}_{-0.03} & -0.07^{+0.06}_{-0.06} & -0.01^{+0.08}_{-0.08} \\
-0.07^{+0.07}_{-0.06} & 0.10^{+0.03}_{-0.03} & 1 & -0.14^{+0.04}_{-0.03} & -0.17^{+0.05}_{-0.05} & -0.04^{+0.08}_{-0.07} \\
-0.04^{+0.06}_{-0.06} & -0.26^{+0.03}_{-0.03} & -0.14^{+0.04}_{-0.03} & 1 & 0.03^{+0.05}_{-0.06} & 0.03^{+0.08}_{-0.08} \\
0.10^{+0.07}_{-0.07} & -0.07^{+0.06}_{-0.06} & -0.17^{+0.05}_{-0.05} & 0.03^{+0.05}_{-0.06} & 1 & 0.06^{+0.08}_{-0.08} \\
-0.00^{+0.08}_{-0.07} & -0.01^{+0.08}_{-0.08} & -0.04^{+0.08}_{-0.07} & 0.03^{+0.08}_{-0.08} & 0.06^{+0.08}_{-0.08} & 1 \\
\end{array}
\end{pmatrix}.
\end{multline}
\color{black}

We find weak correlations between our color-excess parameters and the input features; 
\color{purple}
only the $ -0.17 \pm 0.05$ between $E_{\gamma^0}(B-V)$ and $EW_{Si}$ has significance much larger than $1 \sigma$.
\color{black}
Recall that $EW_{Si}$ is correlated with light-curve shape, which is correlated with host-galaxy (including dust) properties 
\citep{2000AJ....120.1479H, 2003MNRAS.340.1057S}.
An extensive discussion of the correlations between the spectral features of the SNfactory data set can be found in \citet{chotard:thesis}
and \citet{leget:thesis}.

@@

\color{purple}
For brevity, we do not present here the analysis of simulated data used to validate and quantify bias in our results.
The simulation tests in \S\ref{model2:sec} are representative for both Model~1 of this Section and Model~2 of the subsequent section.
\color{black}

@@

%\color{purple}

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.95in]{output11/plane0.pdf}
   \includegraphics[width=2.95in]{output11/plane1.pdf}
%   \includegraphics[width=2.95in]{output11/plane0BVR.pdf}
%   \includegraphics[width=2.95in]{output11/plane1BVR.pdf}
   \caption{
   \color{purple}
   Two perspectives (left and right columns) of the $UVI$-space
of the vectors corresponding to $\gamma^i_X/(\gamma^i_B-\gamma^i_V)$ of our model (solid lines),
   and $a(X)$, $b(X)$ of the \citetalias{1999PASP..111...63F} model (dashed lines).  The combination $a(X)+b(X)/2.44$ is shown in the dotted red
   line: it is almost perfectly superimposed on $\gamma^0_X/(\gamma^0_B-\gamma^0_V)$.   The combination $-7.31a(X)+b(X)$ is shown in the dotted blue
   line: it is
   almost  perfectly superimposed on $\gamma^1_X/(\gamma^1_B-\gamma^1_V)$ in $UVI$.  
   \label{plane:fig}}
\end{figure}

@@

\color{purple}
The preferred values for $\kappa$ can be determined from the matrix elements of Eqn.~\ref{model11trans:eqn}: $\kappa^{-1}_1=2.82/1.15 =2.44$, $\kappa_2=-5.27/0.72=-7.31$.
The directions of the $\gamma$ vectors, and  the expectations from  \citetalias{1999PASP..111...63F} with $R=\kappa^{-1}$, are shown in 
Figure~\ref{plane:fig}.  There is almost perfect alignment between $\gamma^0$ and $R=2.44$ dust.
 From the Monte Carlo chains, we find a 68\% credible interval for $\kappa^{-1}_1$ of 2.26--2.63.

@@

\color{purple}
As a check of our analysis implementation, we generate ten simulated datasets with the same numbers and uncertainties as our data, based on a
SALT2 $x_1=0$, $c=0$ template and  \citetalias{1999PASP..111...63F} dust.  The supernovae
are assigned $E^F(B-V)$ parameters equal to the $B-V$ color excesses determined in our sample
(making an approximative correspondence between our observed top-hat filter colors and the 
model parameter $E^F(B-V)$) , and $R^F$ values drawn from a log-normal
distribution with a mean of $\ln{2.5}$ and 0.05 standard deviation.  The $\alpha$, $\beta$, and $\eta$ parameters are set to zero, so that dust and
measurement uncertainty are the only sources of variation in magnitudes.  We stack the results of the 10 simulations: the input $R^F=2.5$ is centered
within the 68\% credible interval for $\kappa^{-1}_1$ of 2.38--2.69.

@@

Given the association of  our parameter $\gamma^0_V k_0$ with changes in $A^F_V$ of an $R^F =\kappa^{-1}_1$ dust model up to an additive offset,
the former's distribution can be compared to the expectation for the latter.
Figure~\ref{k0_med:fig} shows the histogram of
median values of $\gamma^0_V k_0$ 
\color{purple}
relative to that of an arbitrary supernova  $\gamma^0_V k_0|_0$.
()
\color{black}
The distribution is non-Gaussian, having a sharp rise in the blue and an extended tail in the red.  This is consistent
with the expectation of dust considering the spatial distribution of supernovae within galaxies and the distribution of galaxy orientations with respect to the observer \citep{1998ApJ...502..177H}.  This distribution is given by the data, in that it is independent of any knowledge or prior
for the distribution of dust extinction the SN population is expected to suffer.

Figure~\ref{k0_med:fig} also shows the distribution of $\gamma^1_V k_1$, which can be attributed to 
extinction corrections due to deviations away from the canonical $\kappa_1$ value.   The distribution has a long
negative tail and is peaked at its positive edge, and by construction has a mean of zero.
The implication is that while the bulk of supernovae have positive values associated with $R>1/\kappa_1$, there is a tail
of negative values associated with $R<1/\kappa_1$.  The plot of   $\gamma^0_V k_0$ versus $\gamma^1_V k_1$  shown in Figure~\ref{kk:fig} shows that
the positive edge in $\gamma^1_V k_1$ persists over the full range of  $\gamma^0_V k_0$
%\color{purple}
%While at low  $\gamma^0_V k_0$
%\color{black}
%On the other hand, lower values of $\gamma^1_V k_1$
%occur at larger values of  $\gamma^0_V k_0$.
These findings are consistent with previous results:
\citet{2014ApJ...789...32B, 2015MNRAS.453.3300A} deduce a wide range of dust behavior $1.5<R^F<3$ encountered by the SN~Ia population.
%\citet{2011ApJ...729...55F} and the above authors
%find that supernovae with large extinction 
%preferentially exhibit low $R^F$.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{output11/deltagamma0_med.pdf}
   \includegraphics[width=2.8in]{output11/deltagamma1_med.pdf}
      \caption{
      \color{purple} Normalized ideogram and histogram 
      of
median values of $\gamma^0_V k_0-\gamma^0_V k_0|_0$, which is associated with the $A^F_V$ of an $R^F \sim 1/\kappa_1$ dust model
(left) : $\gamma^1_V k_1-\gamma^1_V k_1|_0$,
which  is associated with extinction corrections due to deviations away from the canonical $\kappa_1$ value (right).
      \color{black}
   \label{k0_med:fig}}
\end{figure}
